---
title: 'Wrangling Basics: Slice, Summarize and Visualization '
author: "Yi Yin"
date: "10/01/2018"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Scope

- Wrangle a data set posted on [U.S. Chronic Disease Indicators (CDI)](https://data.cdc.gov/Chronic-Disease-Indicators/U-S-Chronic-Disease-Indicators-CDI-/g4ie-h725)
- Produce some summary statistics
- Visualize the correlation between binge drinking prevalence and poverty in U.S. States. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.path="images/",
                      cache.path="cache/",
                      cache=TRUE,
                      message=FALSE,
                      warning=FALSE)

```

# Data Wrangling
## Slice
1. Remove the irrelevant columns
2. Select columns related to 
  a) **Binge Drinking**:
_Binge drinking prevalence among adults aged >= 18 years_, Crude Prevalence in Percent. 
We would like to obtain this variable for the overall population, as well separately for _females_ and _males_.
  b) **Poverty**:
_Poverty, Crude Prevalence in Percent_. We only want the overall poverty prevalence to make things a bit easier.
```{r, cache=TRUE, eval=FALSE, echo=TRUE, results='hide'}
cdi <- readr::read_csv("U.S._Chronic_Disease_Indicators__CDI_.csv")
cdi_rm <- cdi%>%dplyr::select(YearStart,
                              YearEnd,
                              LocationAbbr,
                              LocationDesc,
                              Question,
                              DataValue,
                              DataValueType,
                              Stratification1)%>%
        dplyr::filter(Question %in% 
                              c('Binge drinking prevalence among adults aged >= 18 years', 'Poverty') 
                      & DataValueType == 'Crude Prevalence')%>%dplyr::select(-DataValueType)


cdi_rm <- cdi_rm%>%dplyr::filter(Stratification1 %in% c('Overall', 'Male', 'Female' ))
```

## Transform
3. Convert the data set into tidy format via `tidyr` package.
```{r, cache=TRUE, eval=FALSE, echo=TRUE, results='hide'}
cdi_tidy <- cdi_rm%>% unite(category, Question, Stratification1) %>%
        spread(key = category, value = DataValue)

head(cdi_tidy)
```

## Save
4. Rename variables, save cleaned data set.
```{r, cache=TRUE, eval=FALSE, echo=TRUE, results='hide'}
cdi_tidy <- cdi_tidy%>%dplyr::select(-YearStart, -Poverty_Female, -Poverty_Male)%>%
        dplyr::rename(state = LocationDesc,
                      stateabb = LocationAbbr,
                      year = YearEnd,
        binge_all = `Binge drinking prevalence among adults aged >= 18 years_Overall`,
        binge_male = `Binge drinking prevalence among adults aged >= 18 years_Male`,
        binge_female = `Binge drinking prevalence among adults aged >= 18 years_Female`,
        poverty = Poverty_Overall)

# Save the cleaned data set as binge_clean.csv
write.csv(cdi_tidy, "binge_clean.csv")
```


# Summary Statistics 
## Sorting
5. Produce a table that shows the overall, female, and male binge drinking prevalences across U.S. States in the most recent year of data for the Top 10 binge drinking states (i.e. the ones with the highest prevalence in the overall population). 

```{r, cache=TRUE, message=FALSE}
library(tidyverse)
cdi_tidy <- read_csv('binge_clean.csv')
cdi_tidy%>%dplyr::filter(year == 2016 )%>%
        arrange(desc(binge_all))%>%
        dplyr::select(state, binge_all,binge_female,binge_male)%>%
        top_n(10,binge_all)
```

## Regression
6. Make a simple scatter plot showing the correlation between the overall poverty prevalence in a state and the prevalence of binge drinking in the overall population. 

```{r, cache=TRUE, message=FALSE}
cdi_tidy%>%ggplot(aes(x= binge_all, y = poverty, 
                      label = stateabb )) +
        geom_point(size = 0.5) + 
        geom_text(check_overlap = TRUE, hjust = -0.05, nudge_x = 0.05)+
        geom_smooth(size = 0.8, colour = '#4EB6D4')+
        geom_smooth(method='lm', formula =  y ~ x, 
                    se = FALSE, colour = '#D69CBE', linetype = 2)+
        ggtitle("Overall Poverty Prevalence VS. Prevalence of Binge Drinking")+
        xlab("prevalence of binge drinking")+
        ylab("overall poverty prevalence")+
        scale_x_continuous(labels = function(x) paste0(x, "%"))+
        scale_y_continuous(labels = function(y) paste0(y, "%"))+
        theme_bw(base_size = 11, base_family = "Palatino")+
        theme(plot.title = element_text(hjust = 0.5))


```

Overall, there is a slight negative association between the overall poverty prevalence in a state and binge drinking prevalence in the general population. However, the variance in poverty prevalence of extreme high and low binge drinking states are larger than the variance in poverty prevalence of medium binge drinking states.

```{r}
model <- lm( poverty ~ binge_all , data = cdi_tidy)
library(olsrr)
olsrr::ols_test_breusch_pagan(model)

```

By the Breush-Pagan test we can see the heteroscedasticity in `poverty ~ binge_all` model. Therefore, it is hard to draw inference from a simple linear model between this two variable.


## Grouping

7.Calculate the average annual growth rates (in percent) of overall binge drinking across states for the years the data is available.

Provide a table of the 5 states with the largest increases and the 5 states with the largest decreases in binge drinking prevalence over the time period.

```{r}

largest_increases <- cdi_tidy%>%select(state, year, binge_all)%>%
        dplyr::group_by(state)%>%
        dplyr::filter(is.na(binge_all) == FALSE, 
                      state != "United States")%>%
        summarise(avg_growth = (last(x = binge_all, order_by = year) -
                                first(x = binge_all, order_by = year)
                                )/length(unique(year)))%>%
        arrange(desc(avg_growth))%>%
        slice(1:5)


largest_decreases <- cdi_tidy%>%select(state, year, binge_all)%>%
        dplyr::group_by(state)%>%
        dplyr::filter(is.na(binge_all) == FALSE, 
                      state != "United States")%>%
        summarise(avg_growth = (last(x = binge_all, order_by = year) -
                                first(x = binge_all, order_by = year)
                                )/length(unique(year)))%>%
        arrange(avg_growth)%>%
        slice(1:5)

dplyr::union(largest_increases,largest_decreases)%>%
        arrange(desc(avg_growth))
        
        
   

        

```

