---
title: "Wrangling Advance: Build Pipeline"
author: "Yi Yin"
date: "2018/10/16"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: console
---
# Scope
- Build pipeline on World Bank data set 
- Explore the relationship between infant mortality and GPD per capita over time 
- Group data by region/country to compare with the overall regression
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.path="images/",
                      cache.path="cache/",
                      cache=TRUE,
                      message=FALSE,
                      warning=FALSE)

```


```{r, echo=FALSE}
content <- mget(ls("package:base"), inherits = TRUE)
base_functions <- Filter(is.function, content)
rm(content)
library(purrr)
library(dplyr)
```

# Cleaning
## Slice and Rename

Use the [`WDI` package](https://cran.r-project.org/web/packages/WDI/WDI.pdf) to obtain country-year data on infant mortality and GDP per capita from the World Bank database. 

The indicator short names in the World Development Indicators are:  

  - `NY.GDP.PCAP.PP.KD` - GDP per capita, PPP (constant 2005 international)  
  - `SP.DYN.IMRT.IN` - Mortality rate, infant (per 1,000 live births)

Rename two variables to readable name: infant mortality (`infmort`) and GPD per capita (`gdpcap`). 

```{r}
library(WDI)
wdi <- WDI(country = "all", 
    indicator = c("NY.GDP.PCAP.PP.KD", "SP.DYN.IMRT.IN"), 
    start = 1960,
    end = 2017, 
    extra=TRUE)

dat<-wdi%>%dplyr::rename(gdpcap = NY.GDP.PCAP.PP.KD,
                    infmort = SP.DYN.IMRT.IN)%>%
        dplyr::filter(longitude != "", latitude != "")



```

## Sorting

Present a list of countries with the 5 highest and the 5 lowest infant mortality rates in 2017 along with their GDP per capita in that year.

```{r}
# 5 highest
dat%>%filter(year == 2017)%>%
        arrange(desc(infmort))%>%
        filter(between(row_number(), 1, 5))%>%
        select(country, infmort, gdpcap)%>%
        # and 5 lowest
        dplyr::full_join(dat%>%filter(year == 2017)%>%
        arrange(infmort)%>%
        filter(between(row_number(), 1, 5))%>%
        select(country, infmort, gdpcap))


```


## Overall Regression
Run an overall regression predicting infant mortality from GDP per capita. I find a significant negetive association between infant mortality and GDP per capita
```{r}
summary(lm(infmort ~ gdpcap, data = dat))$coefficients
```


### Scatter Plot
```{r}
library(ggplot2)
dat%>%select(infmort, gdpcap)%>%
       ggplot(aes(gdpcap, infmort))+
        geom_point()+
        geom_smooth(method = lm, se = FALSE)+
        ggtitle('Infant mortality by GDP per capita')+
        theme_bw(base_size = 11, base_family = "Palatino")+
        theme(plot.title = element_text(hjust = 0.5))
        
```

## Regression by Region
Let’s see if the overall relationship holds within world regions. 

Split the the data by region and create a nested list that contains these split data frames.

Build a single pipeline, re-estimate the regression for each of these data frames, obtain the coefficient estimates for the slope for each region and present them in a table (sorted by size of the coefficient).


```{r}
dat%>%split(dat$region)%>%
        map(possibly(~ lm(infmort ~ gdpcap, data = .), 
                     otherwise = NA_real_))%>%
        discard(is.na(.) == T) %>%
        map(~ coef(.))%>%
        map_dbl("gdpcap") %>% 
        tibble(region = attr(., "names"), 
               coef = .)%>%
        arrange(coef)
```




## Regression by Country

Let’s also check if the relationship is consistent if estimated within countries. Split the overall data by country and create a nested list that contains these split data frames. Again, estimate a simple linear regression, predicting infant mortality by GDP per capita. 

We find 6.06% the within-country regressions have a positive relationship (at the 95% confidence level) between GDP per capita and infant mortality

```{r}
# extract slope p-value for 95% confidence level
pvalue <- dat%>%split(dat$country)%>%
        map(possibly(~ lm(infmort ~ gdpcap, data = .),
                     otherwise = NA_real_))%>%
        discard(is.na(.) == T) %>%
        map(summary)%>%map_dbl(possibly(
                        ~ .$coefficients['gdpcap', 'Pr(>|t|)'],
                         otherwise = NA_real_)) 
```


```{r}
dat%>%split(dat$country)%>%
        map(possibly(~ lm(infmort ~ gdpcap, data = .),
                     otherwise = NA_real_))%>%
        discard(is.na(.) == T) %>%
        map(summary)%>%map_dbl(possibly(
                        ~ .$coefficients['gdpcap', 'Estimate'],
                         otherwise = NA_real_))%>%
        tibble(country = attr(., "names"), 
                  coef = .)%>%cbind(pvalue)%>%
        filter(pvalue < 0.05)%>%
        mutate(pos_slope = ifelse(coef >0, T, F))%>%
        group_by(pos_slope)%>%summarise(n = n())%>%
        mutate(percent = n/sum(n)* 100)

```


### Histogram of the Slope Coefficients

```{r}
dat%>%split(dat$country)%>%
        map(safely(~ lm(infmort ~ gdpcap, data = .)))%>%
        transpose()%>%pluck("result")%>%discard(is.null)%>%
        map(~ coef(.))%>%
        map_dbl("gdpcap")%>%hist(breaks = 100,
                                 col = "#53B2D6",
                                 border = '#53B2D6',
                                 xlab = "Coefficients", 
                                 main = "Histogram of Slope Coefficients")

```

With regard to the absolute value difference, `Malawi` are most at odds with the overall regression estimated.



```{r}
dat%>%split(dat$country)%>%
        map(possibly(~ lm(infmort ~ gdpcap, data = .),
                     otherwise = NA_real_))%>%
        discard(is.na(.) == T) %>%
        map(summary)%>%
        map_dbl(possibly(
                        ~ .$coefficients['gdpcap', 'Estimate'],
                         otherwise = NA_real_))%>%
        tibble(country = attr(., "names"), 
                  coef = .)%>%cbind(pvalue)%>%
        mutate(diff = abs(coef - (-0.0009353066)))%>%
        arrange(desc(diff))%>%filter(row_number() == 1)
```

With regard to the largest positive slope, `Madagascar` are most at odds with the overall regression estimated.
```{r}
dat%>%split(dat$country)%>%
        map(possibly(~ lm(infmort ~ gdpcap, data = .),
                     otherwise = NA_real_))%>%
        discard(is.na(.) == T) %>%
        map(summary)%>%map_dbl(possibly(
                        ~ .$coefficients['gdpcap', 'Estimate'],
                         otherwise = NA_real_))%>%
        tibble(country = attr(., "names"), 
                  coef = .)%>%cbind(pvalue)%>%
        arrange(desc(coef))%>%filter(row_number() == 1)

```


## Added Squared Term
Estimate regressions of infant mortality on GDP per capita and the square of GDP per capita. Compare the adjusted R-Squareds of the models with and without the added squared term of GDP per capita. 


### Without the Added Squared Term
```{r}

dat%>%split(dat$country)%>%
        map(safely(~ lm(infmort ~ gdpcap, data = .)))%>%
        transpose()%>%pluck("result")%>%discard(is.null)%>%
         map(summary) %>% map_dbl("adj.r.squared")%>%mean()

```

### With the Added Squared Term

The average model adjusted R-squared with the squared term is 0.08 larger than above.
```{r}

dat%>%split(dat$country)%>%
        map(safely(~ lm(infmort ~ I(gdpcap^2)+ gdpcap, data = .)))%>%
        transpose()%>%pluck("result")%>%discard(is.null)%>%
         map(summary) %>% map_dbl("adj.r.squared")%>%mean()


```

I recommend to keep the squared term since the average model adjusted R-squared with the squared term is larger than the without counterpart.


