---
title: "String Advance: Analyze Tweets for NYC Subway Delay Pattern"
author: "Yi Yin"
date: "2018/11/17"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
editor_options: 
  chunk_output_type: console
---
# Scope
- Extract multiple formats of data from string (noun, time, number, etc.)
- Manipulate natural language strings without preset tokenization.
- Visualize and generate insights from text analysis.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.path="images/",
                      cache.path="cache/",
                      cache=TRUE,
                      message=FALSE,
                      warning=FALSE)

```

# Statistical Analysis on Text Data

I'm going to learn the pattern of NYC subway delay from tweets on the New York City subway services' official twitter account.


## Data Slice

```{r}
rm(list = ls())
mta <- read_rds("mta.RDS")

data <-mta%>%filter(is_retweet == FALSE,
                    str_detect(text, "^@.*") == FALSE)%>%
        select(text, created_at)
```

## Text Extraction

Time of Delay

Pattern: XX trains are running with delays because of xx problems.
```{r}
library(lubridate)
library(tidyverse)
timed <- data%>%
        filter(str_detect(text, "trains are running with delays") == TRUE)%>% 
        mutate(w_day = wday(created_at, label = T),
               clock = lubridate::hour(created_at))%>%mutate(period=
    case_when(
      between(clock, 6, 9) ~ "mornings",
       between(clock, 10, 14) ~ "mid-day",
       between(clock, 15, 17) ~ "afternoon",
      between(clock, 18, 21) ~ "evening",
      between(clock, 22, 24) ~ "night",
      between(clock, 0, 5) ~ "night"
    ))

```

## Frequency Visualization

A table that shows delays by day of the week and time of day. Use the following time periods: mornings (6-10), mid-day (10-15), afternoon (15 - 18), evening (18- 22), and night (22 - 24; 0-6). 

### Delays by Day of the Week 
```{r}
timed%>%group_by(w_day)%>%summarise(n=n())

```

```{r, echo=FALSE}
timed%>%group_by(w_day)%>%summarise(n=n())%>%
        ggplot(aes(w_day, n, fill = 1/n))+
        geom_bar(stat="identity")+ 
        xlab('Day of the week')+
        ylab('Number of Delays')+
        theme_bw(base_family = 'Palatino')+
        theme(legend.position="none")

```

### Delays by and Time of Day

```{r}
timed$period <- factor(timed$period, ordered = TRUE, 
                    levels = c("mornings", 
                               "mid-day",
                               "afternoon", 
                               "evening",
                               "night"))

timed%>%group_by(period)%>%summarise(n=n())
```

```{r, echo=FALSE}
timed%>%group_by(period)%>%summarise(n=n())%>%
        ggplot(aes(period,n ,fill = 1/n))+
        geom_bar(stat="identity")+ 
        xlab('Time of day')+
        ylab('Number of Delays')+
        theme_bw(base_family = 'Palatino')+
        theme(legend.position="none")

```

###  Type of Delay
Among the set of tweets in part b), try to categorize the types of delays. No need to be exhaustive but try to pick up the top 3-5 reasons for delays. 
```{r}
typed <- timed$text%>%
        str_split("because of", n= 2, simplify = TRUE)%>%
        .[,2]%>%word(1, 4)%>%
        as_tibble()%>%
        rename(reason = value)



typed%>%filter(is.na(reason) == F )%>%
        group_by(reason)%>%
        summarise(n = n())%>%
        top_n(20, n)%>% arrange(desc(n))
        
        
```

Combine them into reasonable categories if necessary (e.g. signal problems, medical, technical problems, etc.). Provide an overview (table or graph) of which types of delays are most common.

```{r}
typed%>%filter(is.na(reason) == F )%>%
        group_by(reason)%>%
        summarise(n = n())%>%
        top_n(6, n)%>% arrange(desc(n))%>%
        ggplot(aes(reason,n ,fill = 1/n))+
        geom_bar(stat="identity")+ 
        xlab('Reason')+
        ylab('Number of Delays')+
        theme_bw(base_family = 'Palatino')+
        theme(legend.position="none")

```

###  Which train lines affected?

Write a regex pattern that captures which train lines are affected by delays. 
```{r}
lined <- timed%>%select(text, w_day)%>% mutate(train = str_extract(text, 
                                  "(.*)(?= train[s]* are)"))

lined <- lined%>% 
        mutate(train_s = str_remove_all(train,
                                        ".*bound[\\b\\s]"))%>%
        mutate(weekend = ifelse(w_day %in% c("Sun", "Sat"), 
                                "weekend", 
                                "weekday"))


lines <-lined%>% select(train_s, weekend)%>% 
        mutate(line = str_remove_all(train_s, "\\b[A-Za-z]{2,}\\b")%>%
        str_remove_all("[^A-Z0-9]")%>%
        str_replace_all('([A-Z0-9])\\1+', '\\1'))%>%
        select(-train_s)%>%
        mutate(i = row_number())

# max(map_dbl(lines$line, nchar))
# lines$i[nchar(lines$line) == 8]

line_n <- paste0("line", 1:8)

library(reshape2)

lines <- cbind(reshape2::colsplit(lines$line, "", names = line_n),
               lines[,3], lines[,1])

tidy <- lines %>% 
  reshape2::melt(variable.name = "note",
                 value.name = "train",
                 id.vars = c("i", 'weekend'))

```

A summary of which train lines are affected by weekday vs. weekend.

```{r, fig.align='center', fig.height=10, fig.width= 15}

tidy%>%filter(train !="")%>%
        select(train,weekend)%>%
        group_by(train,weekend)%>%summarise(n = n())%>%
        mutate(weekend, percent = paste0(sprintf("%1.1f",
                                          n/sum(n)*100),"%"))%>%
  ggplot(aes(train, n, fill= weekend))+
        geom_bar(stat="identity") +
        geom_text(aes(label=percent, y=n-15), size = 3.8,
                  vjust=0.2)+
        ylab("number of delays")+ xlab("line")+
        ggtitle('Lines affected by Weekday vs. Weekend')+
        theme_bw(base_family = "Palatino", base_size = 13)+
        theme(plot.title = element_text(hjust = 0.5),
              legend.position="bottom")
```

The graph above shows us weekdays have a higher chance for NYC subway delays, but this is inconsistent with my personal experience: I encountered delay almost every weekend.

Then I realized that there are five days counted as weekdays while only two days on weekends. So we need to normalize the probability.

```{r,fig.height=10, fig.width= 12}
tidy%>%filter(train !="")%>%
        select(train,weekend)%>%
        group_by(train,weekend)%>%summarise(n = n())%>%
        mutate(probability =  paste0(sprintf("%1.1f",
                                             case_when(weekend=="weekday" ~ n/sum(n)*500/7, 
                                       weekend=="weekend" ~ n/sum(n)*200/7)),"%"))%>%
        ggplot(aes(train, n, fill= weekend))+
        geom_bar(stat="identity",position=position_dodge()) +
      geom_text(aes(label=probability, y=n+2), size = 3.2,
                  vjust=0.1)+
        ylab("number of delays")+ xlab("line")+
        ggtitle('Lines affected by Weekday vs. Weekend\n(Normalized Probability)')+
        theme_bw(base_family = "Palatino", base_size = 13)+
        theme(plot.title = element_text(hjust = 0.5),
              legend.position="bottom")+
        scale_fill_brewer(palette="Accent")
```


Alas, still, there is a much higher chance of delay during weekdays. Then I realized I lived in the upper west side near Columbia University as a graduate student. On weekdays, I seldom ride on the metro, so I don't retain subway delays on weekdays. But on weekends, I take subways to midtown and downtown and always come across line cancellation, tunnel construction, routine changes, etc. 😂  Such an excellent example on data analysis triumphs anecdotes and biased experiences.